Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-03-17 Grassmannian_cells.tex
Intro to section 4

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest that the dimensions of cells can be computed using this combinatorics (or the equivalent combinatorics of compatible pairs).
This is made precise in Conjecture~\ref{conj:cell dimensions}.
%\sayT{maybe we can erase some parts which show up twice in the introduction?}
\subsubsection*{Acknowledgements}We would like to thank Giovanni Cerulli Irelli, Hans Franzen, Oliver Lorscheid and Markus Reineke for very fruitful discussions related to this project.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

The functor $F_Q$ induces a map $F_Q:\ZZ^{\hat Q_1}\to \ZZ^{Q_1}$.
We say that a dimension vector $\hat \bfe$ of $\hat Q$ is \emph{compatible} with $\bfe$ if $F_Q(\hat\bfe)=\bfe$.
The group $A_Q$ acts on $\hat Q$ via translation, this induces actions of $A_Q$ on $\rep\hat Q$ and on $\ZZ^{\hat Q_1}$.
%We say two representation (resp. dimension vectors) are \emph{equivalent} if they can be related by the action of $A_Q$.
The analogous observation can also be made for $\widetilde Q$.
If $X$ is a representation of $\hat Q$ (resp. $\widetilde Q$), we denote by $X_\chi$ (resp. $X_w$) the representation obtained via translation by $\chi\in A_Q$ (resp. $w\in W_Q$).

@@

In this section, we aim to establish the existence of cell decompositions for quiver Grassmannians of (truncated) preprojective representations of $K(n)$ and its universal covering quiver.
By a \emph{cell decomposition} of an algebraic variety $X$, we mean a filtration $\varnothing=X_{k+1}\subset X_k\subset \cdots \subset X_2\subset X_1=X$ of $X$ by closed subsets $X_i\subset X$ so that each $X_i\setminus X_{i+1}$ is isomorphic to an affine space.
Alternatively, a cell decomposition of $X$ is a collection of disjoint subsets $U_1,\ldots,U_k\subset X$, each isomorphic to an affine space, such that each $X_i=U_i\sqcup U_{i+1}\sqcup\cdots\sqcup U_k$ is closed in $X$ with $X_1=X$.
We call the subsets $U_i\subset X$ the \emph{affine cells} for this cell decomposition.

@@

  Set $c_{\alpha_1}=1$ and choose $c_{\alpha_l}$ recursively in such way that 
  \[c_{\alpha_l}\geq 2(K+1)\sum_{k=1}^{l-1} c_{\alpha_k}\]
  for $l=2,\ldots,n$.  
  Then let $f:A_Q\to\ZZ$ be the group homomorphism defined by $f(e_\alpha)=c_\alpha$ for all $\alpha\in Q_1$ and define $d(i,\chi):=f(\chi)$ for $(i,\chi)\in\supp(\hat X)$.
  %In particular, for $(i',\chi)$ with $\chi=\sum_{l=1}^n \kappa_l e_{\alpha_l}$, we have
  %\[d(i',\chi)=\sum_{l=1}^n \kappa_l c_{\alpha_l}.\]

@@

  \sayD{When $\Psi_{m+1,I,j}$ is surjective, $\cG_{\tbff,\tbfg}$ is a closed subset of the Grassmannian.
  The map is not surjective exactly when $\tbfg=\udim \tilde P_{m+1}^I$ and so we get a cell decomposition by taking the affine cells from this $\cG_{\tbff,\tbfg}$ last.}

@@

 Proposition~\ref{fibers} shows that the fiber of $\Psi_{\tbff,\tbfg}$ over $(U,V)\in B_\ell\times C_k$ is empty exactly when $\tbfg=\udim\tilde P_{m+1}^I$ and one of the following conditions is satisfied
  \begin{itemize}
    \item $m=2$ with $\tbff_{(2,\alpha_j^{-1})}\ne0$ or $\tbff_{(1,e)}\ne0$;
    \item $m\ge3$, with $\pi_{m,j}^\bfJ(U)=0$ for all $U\in B_\ell$.
  \end{itemize}
By induction, the compatibility condition ($\dagger$) is true for $\pi_{m,j}^\bfJ$  which shows that either all or non fibres over $B_l\times C_k$ are empty. This shows the compatibility with the image. If the fibre is not empty, Theorem \ref{vb} gives that we obtain an affine cell in $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$  of the form $B_\ell\times C_k\times\AA^d$ with~$d=\langle\tbfg,\udim\tilde P_{m,j}-\tbff\rangle$. Altogether this establishes a cell decomposition of $\Gr_{\tbfe}^{\tilde Q}(\tilde P_{m+1}^J)$.

  For $V\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ which is contained in such a cell, we have $\pi_{m+1}^{J,I}(V)\in C_k$.
  But then $\pi_{m+1}^{J,\bfI}(V)\ne0$ if and only if $\pi_{m+1}^{I,\bfI}(W)\ne0$ for all $W\in C_k$ and thus $\pi_{m+1}^{J,\bfI}(V')\ne0$ for all $V'$ in this cell $B_\ell\times C_k\times\AA^d$ of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$. This shows ($\dagger$) for $\pi_{m+1}^{J,\bfI}$.

\end{proof}

%Note that we can also consider 
%$\Psi_1:\Gr^{\tilde Q}_\tbfe(\tilde P^{(k-1)}_{m})\to\bigsqcup_{\tbff+\tbfg=\tbfe} \Gr^{\tilde Q}_\tbff(\bigoplus_{i=1}^{k}\tilde P_{m-1,i})\times \Gr^{\tilde Q}_\tbfg(\tilde P_m^{(k)})$ where $\tilde P_m^{(n-1)}=\tilde P_{m-1}^{(1)}$ (up to renumbering the arrows). But then one needs to adjust $\Psi_2$ because the factor of $\bigoplus_{i=1}^{k}\tilde P_{m-1,i}$ were $(\tau\tilde P_m^{(k)})^k$.

>>>>

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest that the dimensions of cells can be computed using this combinatorics (or the equivalent combinatorics of compatible pairs).
This is made precise in Conjecture~\ref{conj:cell dimensions}.
\subsubsection*{Acknowledgements}We would like to thank Giovanni Cerulli Irelli, Hans Franzen, Oliver Lorscheid and Markus Reineke for very fruitful discussions related to this project.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

The functor $F_Q$ induces a map $F_Q:\ZZ^{\hat Q_1}\to \ZZ^{Q_1}$.
We say that a dimension vector $\hat \bfe$ of $\hat Q$ is \emph{compatible} with $\bfe$ if $F_Q(\hat\bfe)=\bfe$.
The group $A_Q$ acts on $\hat Q$ via translation, this induces actions of $A_Q$ on $\rep\hat Q$ and on $\ZZ^{\hat Q_1}$.
The analogous observation can also be made for $\widetilde Q$.
If $X$ is a representation of $\hat Q$ (resp. $\widetilde Q$), we denote by $X_\chi$ (resp. $X_w$) the representation obtained via translation by $\chi\in A_Q$ (resp. $w\in W_Q$).

@@

In this section, we aim to establish the existence of cell decompositions for quiver Grassmannians of (truncated) preprojective representations of $K(n)$ and its universal covering quiver.
By a \emph{cell decomposition} of an algebraic variety $X$, we mean a filtration $\varnothing=X_{k+1}\subset X_k\subset \cdots \subset X_2\subset X_1=X$ of $X$ by closed subsets $X_i\subset X$ so that each $X_i\setminus X_{i+1}$ is isomorphic to an affine space.
Alternatively, a cell decomposition of $X$ is a collection of disjoint subsets $U_1,\ldots,U_k\subset X$, each isomorphic to an affine space, such that each $X_i=U_i\sqcup U_{i+1}\sqcup\cdots\sqcup U_k$ is closed in $X$ with $X_1=X$.
We call the subsets $U_i\subset X$ the \emph{affine cells} for this cell decomposition.
Given varieties $X$ and $Y$ each with cell decompositions, we may choose an ordering on products of their affine cells (e.g.\ lexicographic) to get a cell decomposition of $X\times Y$. 
Given a variety $X$ with a cell decomposition, we call a subvariety $U\subset X$ \emph{compatible} with the cell decomposition if $U$ can be written as the union of a subset of the affine cells for $X$.
In this case, $U$ also has a cell decomposition given by taking exactly those affine cells which are contained in $U$.

@@

  Set $c_{\alpha_1}=1$ and choose $c_{\alpha_l}$ recursively in such way that 
  \[c_{\alpha_l}\geq 2(K+1)\sum_{k=1}^{l-1} c_{\alpha_k}\]
  for $l=2,\ldots,n$.  
  Then let $f:A_Q\to\ZZ$ be the group homomorphism defined by $f(e_\alpha)=c_\alpha$ for all $\alpha\in Q_1$ and define $d(i,\chi):=f(\chi)$ for $(i,\chi)\in\supp(\hat X)$.

@@

@@

  Proposition~\ref{fibers} shows that the fiber of $\Psi_{\tbff,\tbfg}$ over $(U,V)\in B_\ell\times C_k$ is empty exactly when $\tbfg=\udim\tilde P_{m+1}^I$ and one of the following conditions is satisfied
  \begin{itemize}
    \item $m=2$ with $\tbff_{(2,\alpha_j^{-1})}\ne0$ or $\tbff_{(1,e)}\ne0$;
    \item $m\ge3$, with $\pi_{m,j}^\bfJ(U)=0$ for all $U\in B_\ell$.
  \end{itemize}
  By induction, the compatibility condition ($\dagger$) is true for $\pi_{m,j}^\bfJ$  which shows that either all or non fibres over $B_l\times C_k$ are empty.
  This shows the compatibility with the image.
  If the fiber is not empty, Theorem \ref{vb} gives that we obtain an affine cell in $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ of the form $B_\ell\times C_k\times\AA^d$ with~$d=\langle\tbfg,\udim\tilde P_{m,j}-\tbff\rangle$.
  Altogether this establishes a cell decomposition of $\Gr_{\tbfe}^{\tilde Q}(\tilde P_{m+1}^J)$.

  For $V\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ which is contained in such a cell, we have $\pi_{m+1}^{J,I}(V)\in C_k$.
  But then $\pi_{m+1}^{J,\bfI}(V)\ne0$ if and only if $\pi_{m+1}^{I,\bfI}(W)\ne0$ for all $W\in C_k$ and thus $\pi_{m+1}^{J,\bfI}(V')\ne0$ for all $V'$ in this cell $B_\ell\times C_k\times\AA^d$ of $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$. This shows ($\dagger$) for $\pi_{m+1}^{J,\bfI}$.
\end{proof}
